### 2022-12-09 SALT-abstract/SALT-abstract.tex
updated SALT abstract

@@

\vspace*{-.6\baselineskip}
\noindent 
	We present experimental evidence that \textbf{(i)} the projection of the content of attitude complements varies across entailment-canceling operators (negation, question, modal, conditional) and \textbf{(ii)} this by-operator variation differs across attitude predicates. Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}). Instead, the observed by-operator variation groups predicates by lexical semantic/pragmatic properties that raise important questions for future research on projection.

@@

	There has been, to date, one investigation of by-operator variation: \citet{smith_relationship_2014}, who investigated by-operator variation (negation, conditional) for projective content, observed by-operator variation for some contents (e.g., that of \emph{know}, but not that of clefts) and that this by-operator variation differs across contents. E.g., the content of NRRCs was more projective under conditionals than negation, the opposite pattern was observed for \emph{know, win} and epithets. However, the response task used asked participants to rate how surprised they would be to learn the content under investigation after observing the utterance. Since many factors other than speaker commitments may influence surprisal, it is not clear, whether they (only) measured projection.

@@

\noindent
{\bf Results.}
	Fig.~\ref{fig:figure1} shows the mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean). We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than from under conditionals and modals, the CC of \emph{know} projects less from under negation than under questions, but more from under negation than from under modals, and the CC of \emph{discover} projects less from negation than from under conditionals and questions, and more from under negation than modals. These results (which are supported by linear mixed effects models, see Table \ref{t:models}) are not aligned with the distinction between factive and semi-factive predicates proposed in \citealt{karttunen_observations_1971} or assumed in \citealt{djarv_cognitive_2018}: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators. For {\em know}, we observe a third pattern -- one that is not aligned with a classification as a factive nor as a semi-factive predicate.

@@

\noindent
{\bf Discussion and Conclusion.} 
	Our results clearly show that different entailment-cancelling operators affect projection in different ways, and that the effect of entailment-cancelling operator on projection differs for different verbs. The picture that emerges could not be characterized in terms of Karttunen's characterization of (semi)factives. Our results indicate a need for a more nuanced view of the family-of-sentences diagnostics of projection, as well as a better understanding of the effect of the interaction between lexical semantics and operators on projection.

	Further, we were able to replicate the results of \citealp{degen_are_2022}), observing that projection varies by predicate (e.g., the CC of \emph{be annoyed} is more projective than that of \emph{discover}, and finding no categorical distinction between factive and non-factive predicates either.

	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns. \textbf{Figure \ref{fig:figure2}} gives the mean certainty ratings for the four operators by verb, for some groups of our verbs that group together in the way that they show by-operator variation. We highlight four \lq projection-profiles\rq here: the verbs \emph{pretend, think} are the only verbs that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}). \emph{annouce, confess, admit,} and \emph{reveal} are most projective under conditionals, while there is also a tendency that there is more projection from questions compared to modals and negation, a difference that may not be robust for \emph{announce} (\texttt{C} $>$ \texttt{Q} \textcolor{gray!40}{$>_?$} \texttt{M, N}). \emph{prove, confirm,} and \emph{establish} are more projective under modals and conditionals than under questions and negation (\texttt{M, C} $>$ \texttt{Q, N}). Finally, \emph{inform} and \emph{know} are most projective under questions, and least projective under modals (\texttt{Q} $>$ \texttt{N, Q} $>$ \texttt{M}).

@@

\begin{table}[h]
	\centering
	\vspace{-.7\baselineskip}
	\scalebox{.85}{
		\begin{tabular}{llrrrr}
			Model & & Estimate & Std. Error & t-value\\
			\midrule
			\#1 & Intercept: \emph{be annoyed}/negation & 0.87 & 0.01  & 75.8 & *** \\
			& operator: conditional & -0.12 & 0.02  & -7.38 & *** \\
			& operator: modal & -0.16 & 0.02  & -10.04 & *** \\
			& operator: question & 0.02 & 0.01  & 1.74 & n.s. \\
			\midrule
			\#2 & Intercept: \emph{know}/negation & 0.79 & 0.01 &  69.24 & *** \\
			&		operator: conditional & -0.001 & 0.02  & -0.08 & n.s. \\
			&		operator: modal  & -0.14 & 0.02  & -9.2 & *** \\
			&		operator: question & 0.08 & 0.01  & 5.72 & *** \\
			\midrule
			\#3 & Intercept: \emph{discover}/negation & 0.68 & 0.01  & 59.48 & *** \\
			& operator: conditional & 0.11 & 0.02  & 7.11 & *** \\
			&		operator: modal & -0.06 & 0.02  & -3.6 & *** \\
			&		operator: question & 0.1 & 0.01  & 7.07 & *** \\

@@

\begin{figure}[h]
	\hspace{-2em}
	\begin{minipage}{1.1\textwidth}\centering
		\includegraphics[width=.24\linewidth]{graphs/profile1.pdf}
		\includegraphics[width=.24\linewidth]{graphs/profile2.pdf}
		% \includegraphics[width=.32\linewidth]{graphs/profile3.pdf}\\
		\includegraphics[width=.24\linewidth]{graphs/profile4.pdf}
		\includegraphics[width=.24\linewidth]{graphs/profile5.pdf}
	\end{minipage}
	\vspace{-\baselineskip}
	\caption{Mean certainty-ratings by operator for each verb, showing different \lq verb-profiles\rq.}
	\label{fig:figure2}
\end{figure}

>>>>

\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyear{\#1}}
\newcommand{\citeposs}[1]{\citeauthor{#1}'s}
\newcommand{\citetpos}[1]{\citeauthor{#1}'s (\citeyear{\#1})}

@@

\vspace*{-.8\baselineskip}
\noindent 
	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements varies across four entailment-canceling operators (negation, question, modal, antecedent of conditional) and \textbf{(ii)} that this by-operator variation differs across attitude predicates. Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}). Instead, the observed by-operator variation groups predicates by lexical semantic/pragmatic properties that raise important questions for future research on projection.

@@

	There has been, to date, one investigation of by-operator variation: \citet{smith_relationship_2014}, who investigated by-operator projection variation for negation and conditional embeddings, observed that there is by-operator variation for some contents (e.g., that of \emph{know}, but not that of clefts) and that this by-operator variation differs across contents: The content of NRRCs and \emph{win} was more projective under conditionals than negation, the opposite pattern was observed for \emph{know} and epithets. However, the response task they used asked participants to rate how surprised they would be to learn the content under investigation after observing the utterance. Since factors other than speaker commitment modulate surprisal, it is not clear that they (only) measured projection.

@@

\noindent
{\bf Results.}
	Fig.\ref{fig:figure1} shows the mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean). We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than from under conditionals and modals, the CC of \emph{know} projects less from under negation than under questions, but more from under negation than from under modals, and the CC of \emph{discover} projects less from negation than from under conditionals and questions, and more from under negation than modals. These results (which are supported by linear mixed effects models, see Table \ref{t:models}) are not aligned with the distinction between factive and semi-factive predicates proposed in \citealt{karttunen_observations_1971} or assumed in \citealt{djarv_cognitive_2018}: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators. For {\em know}, we observe a third pattern, one that is neither that of a factive nor that of a semi-factive predicate. Similar considerations apply to \emph{see} and \emph{reveal}, which are also considered (semi-)factive in the literature.

@@

\noindent
{\bf Discussion:  Empirical and theoretical implications.} 
Our results --- that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection --- have several empirical and theoretical implications. First, the results for the predicates typically considered (semi-)factive, namely \emph{be annoyed, know, see, discover} and \emph{reveal}, call into question the assumed distinction between factive and semi-factive predicates (see also \citealt{beaver_have_2010}). Future research that appeals to these two categories must clarify their definition. Second, claims about projection variability must be relativized to the entailment-canceling operator. While our data replicate the result from \citealt{tonhauser_how_2018} that, in a polar question, the CC of \emph{discover} is less projective than that of \emph{know}, this result does not carry over to conditionals. Finally, our results provide further support (from embedding under negation, modals, and conditionals) for the result of \citealt{degen_are_2022}, that projection does not categorically distinguish between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.

\noindent
{\bf Discussion: Novel research question.} Can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how? This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns. The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of Fig.\ref{fig:figure2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality contributed by these predicates. For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in the antecedent of a conditional: This {\bf `Conditional high'} pattern, shown in panel (b), may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates. The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf Negation low'} pattern, shown in panel (c): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not operators), to result in lower projection ratings under negation. Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern, shown in panel (d). The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.

	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns. Fig.~\ref{fig:figure2} gives the mean certainty ratings for the four operators by verb, for some groups of our verbs that group together in the way that they show by-operator variation. We highlight four \lq projection-profiles\rq here: the verbs \emph{pretend, think} are the only verbs that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}). \emph{annouce, confess, admit,} and \emph{reveal} are most projective under conditionals, while there is also a tendency that there is more projection from questions compared to modals and negation, a difference that may not be robust for \emph{announce} (\texttt{C} $>$ \texttt{Q} \textcolor{gray!40}{$>_?$} \texttt{M, N}). \emph{prove, confirm,} and \emph{establish} are more projective under modals and conditionals than under questions and negation (\texttt{M, C} $>$ \texttt{Q, N}). Finally, \emph{inform} and \emph{know} are most projective under questions, and least projective under modals (\texttt{Q} $>$ \texttt{N, Q} $>$ \texttt{M}).

@@

\begin{table}[h]
	\centering
	\vspace{-.7\baselineskip}
	\scalebox{.85}{
		\begin{tabular}{llrrrr}
			Model & & Estimate & Std. Error & t-value\\
			\midrule
			\#1 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01  & 75.8 & *** \\
			& operator: conditional & -0.12 & 0.02  & -7.38 & *** \\
			& operator: modal & -0.16 & 0.02  & -10.04 & *** \\
			& operator: question & 0.02 & 0.01  & 1.74 & n.s. \\
			\midrule
			\#2 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 &  69.24 & *** \\
			&		operator: conditional & -0.001 & 0.02  & -0.08 & n.s. \\
			&		operator: modal  & -0.14 & 0.02  & -9.2 & *** \\
			&		operator: question & 0.08 & 0.01  & 5.72 & *** \\
			\midrule
			\#3 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01  & 59.48 & *** \\
			& operator: conditional & 0.11 & 0.02  & 7.11 & *** \\
			&		operator: modal & -0.06 & 0.02  & -3.6 & *** \\
			&		operator: question & 0.1 & 0.01  & 7.07 & *** \\

@@

\begin{figure}[h]
	\hspace{-2em}
	\begin{minipage}{1.1\textwidth}\centering
		\includegraphics[width=.24\linewidth]{graphs/profile1.pdf}
		\includegraphics[width=.24\linewidth]{graphs/profile2.pdf}
		% \includegraphics[width=.32\linewidth]{graphs/profile3.pdf}\\
		\includegraphics[width=.24\linewidth]{graphs/profile4.pdf}
		\includegraphics[width=.24\linewidth]{graphs/profile5.pdf}
	\end{minipage}
	\vspace{-\baselineskip}
	\caption{Mean certainty-ratings by operator for four predicate patterns in our data.}
	\label{fig:figure2}
\end{figure}
